Challenges to transboundary fisheries management in North America under climate change

### 2020-10-15 Ecol/Manuscript_Draft.Rmd
typos

```{r Table1, eval = T, echo = F}

@@

The results of the present study suggest that climate change will alter the MCP of jointly managed transboundary fish stocks in North America consequently altering Canada’s and the US’s species’ stock-share ratio, regardless of the climate change scenario. These results are aligned with regional [@Morley:2018fn] projections suggesting that climate change will push marine species towards the poles and deeper water [@Pinsky:2013jo] in search of their ecological niche [@Poloczanska:2016kk]. Moreover, IPHC data[^IPHC] suggest that some of these shits are already happening. For example, since 2010, the distribution proportion of Pacific halibut has increased from 9% to 11% in region 2B, from 7.5% to 13% in region 2C, and from 12.3% to 13.5% in region 4CDE. On the other hand, regions 3A and 3B have seen the largest decreases in the IPHC regulatory areas since 2010, from 35.3% to 30.6% and 20.6% to 15.9%, respectively. Similarly, in the Gulf of Maine, the projected stock-share gain of yellowtail flounder and haddock by the US (Figure \ref{fig:Figure5-GOMA-prop}) follows a historical trend where in 2019, Canada's stock-share decreased from 35% to 32% and 60% to 40% relative to 2010, respectively [@Trinko:2019ou].

@@

Geographic barriers [@Cheung:2015fm; @Kleisner:2016je], local temperature gradients [@Pinsky:2013jo], species interactions and human activities [@Serpetti:2017he] might change the rate and direction of species shifts. For the IPHC, geographic barriers might induce a westward increase of stock-share in IPHC regions where species can only migrate northward into the Arctic Ocean through the Bering Sea and Bering Strait [@Cheung:2015fm] (Figure \ref{fig:Figure3-IPHC-prop}). In the Gulf of Maine, future projections could be a response to a temperature gradient shift combined with geographic barriers as southern waters are deeper and warming slower than northern waters according to the ESMs (Figure \ref{fig:FigureS22}). Moreover, Maine has seen its landings of yellowtail flounder increased at the expenses of southern states [@Pinsky:2012kq]. This could be influencing the US gain in MCP in the GoMA in relation to Canada as species shift their distribution from lower latitudes naturally reaching the US (lower) region first. As the effects of climate change endure, even with high mitigation, joint plans should prepare to face changes in the stock-share ratio of transboundary stocks along both coast of North America.

The shifts in the distribution of transboundary stocks can jeopardize management objectives such as conservation measures and gear operation. Fish moving out of fishing grounds and into protected areas could result in a pressure increase to open such area to fishing. Moreover, overlapping shifting stocks could interfere in gear-limitation management rules of multiple fisheries generating conflicts between fleets [@VanDerVoo:2016uc]. The effectiveness of the IPHC-Closed Area ("CA" in Figure \ref{fig:Figure2-IPHC}) in terms of protecting juveniles has been historically questioned as trawling for other species is still allowed in the area [@Karim:2010ut; @IPHC:2017CA]. In 2015, for example, 97% of the trawl by-catch in areas 4CDE and the Closed Area were juveniles [@IPHC:2017CA]. Consequently, the Alaskan trawl fisheries has been closed before reaching annual quota due to the attainment of Pacific halibut bycatch quota limits [@Karim:2010ut]. Thus, the commission has been asked to open the closed area for Pacific halibut fishing, under the premise that the expansion of the trawl fishery is likely reducing any conservation goal for juvenile Pacific halibut [@IPHC:2017CA]. Although not assessed in this study, some trawling target species like Pacific cod (*Gadus macrocephalus*), flathead sole (*Hippoglossoides elassodon*), and Alaskan plaice (*Pleuronectes quadrituberculatus*) have already shift their distributions due to warming waters [@Stram:2009xc] and some are expected to continue shifting in similar direction than Pacific halibut [@Pinsky:2013jo]. The overlap of target species could be addressed by applying ecosystem-based management and dynamic management tools [@Hazen:2018fa] to manage these fisheries and reduce potential loss of sustainable harvest for both the halibut and the trawl fisheries.

@@

<!-- Poleward shifts in Eastern Atlantic off the coast of the US include American lobster (*Homarus americanus*), summer (*Paralichthys dentatus*) and flounders, and red hake (*Urophycis chuss*) from North Carolina to Maine [@Pinsky:2012kq]. -->
<!-- Species such as haddock in the Atlantic [@Morley:2018fn], and big skate (*Raja binoculata*) in the Pacific coast [@Pinsky:2013jo] have been projected to shift away from the poles. -->

>>>>

```{r Table1, eval = T, echo = F, warning=F, message=F, fig.align="H"}

@@

The results of the present study suggest that climate change will alter the MCP of jointly managed transboundary fish stocks in North America consequently altering Canada’s and the US’s species’ stock-share ratio, regardless of the climate change scenario. These results are aligned with regional [@Morley:2018fn] projections suggesting that climate change will push marine species towards the poles and deeper water [@Pinsky:2013ab] in search of their ecological niche [@Poloczanska:2016kk]. Moreover, IPHC data[^IPHC] suggest that some of these shits are already happening. For example, since 2010, the distribution proportion of Pacific halibut has increased from 9% to 11% in region 2B, from 7.5% to 13% in region 2C, and from 12.3% to 13.5% in region 4CDE. On the other hand, regions 3A and 3B have seen the largest decreases in the IPHC regulatory areas since 2010, from 35.3% to 30.6% and 20.6% to 15.9%, respectively. Similarly, in the Gulf of Maine, the projected stock-share gain of yellowtail flounder and haddock by the US (Figure \ref{fig:Figure5-GOMA-prop}) follows a historical trend where in 2019, Canada's stock-share decreased from 35% to 32% and 60% to 40% relative to 2010, respectively [@Trinko:2019ou].

@@

Geographic barriers [@Cheung:2015fm; @Kleisner:2016je], local temperature gradients [@Pinsky:2013ab], species interactions and human activities [@Serpetti:2017he] might change the rate and direction of species shifts. For the IPHC, geographic barriers might induce a westward increase of stock-share in IPHC regions where species can only migrate northward into the Arctic Ocean through the Bering Sea and Bering Strait [@Cheung:2015fm] (Figure \ref{fig:Figure3-IPHC-prop}). In the Gulf of Maine, future projections could be a response to a temperature gradient shift combined with geographic barriers as southern waters are deeper and warming slower than northern waters according to the ESMs (Figure \ref{fig:FigureS22}). Moreover, Maine has seen its landings of yellowtail flounder increased at the expenses of southern states [@Pinsky:2012kq]. This could be influencing the US gain in MCP in the GoMA in relation to Canada as species shift their distribution from lower latitudes naturally reaching the US (lower) region first. As the effects of climate change endure, even with high mitigation, joint plans should prepare to face changes in the stock-share ratio of transboundary stocks along both coast of North America.

The shifts in the distribution of transboundary stocks can jeopardize management objectives such as conservation measures and gear operation. Fish moving out of fishing grounds and into protected areas could result in a pressure increase to open such area to fishing. Moreover, overlapping shifting stocks could interfere in gear-limitation management rules of multiple fisheries generating conflicts between fleets [@VanDerVoo:2016uc]. The effectiveness of the IPHC-Closed Area ("CA" in Figure \ref{fig:Figure2-IPHC}) in terms of protecting juveniles has been historically questioned as trawling for other species is still allowed in the area [@Karim:2010ut; @IPHC:2017CA]. In 2015, for example, 97% of the trawl by-catch in areas 4CDE and the Closed Area were juveniles [@IPHC:2017CA]. Consequently, the Alaskan trawl fisheries has been closed before reaching annual quota due to the attainment of Pacific halibut bycatch quota limits [@Karim:2010ut]. Thus, the commission has been asked to open the closed area for Pacific halibut fishing, under the premise that the expansion of the trawl fishery is likely reducing any conservation goal for juvenile Pacific halibut [@IPHC:2017CA]. Although not assessed in this study, some trawling target species like Pacific cod (*Gadus macrocephalus*), flathead sole (*Hippoglossoides elassodon*), and Alaskan plaice (*Pleuronectes quadrituberculatus*) have already shift their distributions due to warming waters [@Stram:2009xc] and some are expected to continue shifting in similar direction than Pacific halibut [@Pinsky:2013ab]. The overlap of target species could be addressed by applying ecosystem-based management and dynamic management tools [@Hazen:2018fa] to manage these fisheries and reduce potential loss of sustainable harvest for both the halibut and the trawl fisheries.

@@

<!-- Poleward shifts in Eastern Atlantic off the coast of the US include American lobster (*Homarus americanus*), summer (*Paralichthys dentatus*) and flounders, and red hake (*Urophycis chuss*) from North Carolina to Maine [@Pinsky:2012kq]. -->
<!-- Species such as haddock in the Atlantic [@Morley:2018fn], and big skate (*Raja binoculata*) in the Pacific coast [@Pinsky:2013ab] have been projected to shift away from the poles. -->
